# Homework 03

You will analyze data looking at the relationship between **green reputation** and three personality traits--**compassion**, **intellectual curiosity**, and **openness to experiences**.
The dataset includes data from **students** and **non-students**.

```{r}
dictionary <- readr::read_csv("green_dictionary.csv")
green_data <- readr::read_csv("green_data.csv")

library(dplyr)
library(ggplot2)
library(tidyverse)
library(psychTools)
library(patchwork)
library(ggdist)
```

```{r}
view(dictionary)
view(green_data)
```
 
 ## 1. Inspect the item responses (e.g., with graphs or by summarizing distinct values). Is anything unusual?

```{r inspection}
green_data %>% 
  summary()
#What's unusual here is that minimum values are "-99" for several of these items. I'm going to change the "-99" values to "NA."
```
```{r}
green_data <- green_data %>% 
  mutate(across(everything(), 
                ~ na_if(.x, -99))) #changing green_data's -99 values to NA

view(green_data)
```
## 2. Compute total scores for the four scales. Recode variables as needed.

```{r}
reversed <- dictionary %>% 
  filter(Keying < 0) %>% 
  pull(Item)
#making a new variable that filters those items that have a Keying less than 0 out of the dictionary. This will give me all the reverse coded variables, and I should get 14 of those. 

view(reversed)
class(reversed)
#looks good, now I have a new variable with all the items I need to reverse code. 

#now I need to actually reverse code them
# here I'm changing all of my "reversed" values in green_data. changing 1 values to 5 values and so on. the names of the columns will be whatever they were before with an additional "_r" at the end.
green_data_r <- green_data %>% 
  mutate(
    across(all_of(reversed),
           ~ recode(.x, "1" = 5, "2" = 4, "3" = 3, "4" = 4, "5" = 1),
           .names = "{.col}_r")
  )

str(green_data_r)
view(green_data_r)
#I guess I should move forward without using the the original items and only use the recoded ones
green_data_r <- select(green_data_r, -c("green2", "comp6", "comp7", "comp8", "comp9", "comp10", "intel7", "intel8", "intel9", "intel10", "open7", "open8", "open9", "open10"))

view(green_data_r) #ok, now I only have the recoded versions



green_data_totals <- green_data_r %>% 
  rowwise() %>% 
  mutate(greenTotal = mean(c_across(c(green1, green2_r, green3:green5)), na.rm = TRUE),
          compTotal = mean(c_across(c(comp1:comp5, comp6_r:comp10_r)), na.rm = TRUE), 
          intTotal = mean(c_across(c(intel1:intel6, intel7_r:intel10_r)), na.rm = TRUE),
          openTotal = mean(c_across(c(open1:open6, open7_r:open10_r)), na.rm = TRUE)) %>% 
  select(id, student, greenTotal, compTotal, intTotal, openTotal) %>% 
  ungroup()




```

## 3. Rescale the variables so that they go from 0-100 instead of the original range.
     Name the rescaled variables `*_pomp`.
```{r} 
pomp <- green_data_r %>% 
  mutate(across(c(green1:open10_r), 
                ~recode(.x, "1" = 0, "2" = 25, "3" = 50, "4" = 75, "5" = 100), 
                .names = "{.col}_pomp", 
                .before = "greenTotal")) 

#I'm not getting an error message, but the .before argument seems to have no effect. I don't know what's wrong here but it's not a big deal, I think I can just select()

pomp <- select(pomp, green1_pomp:open10_r_pomp)

#ok, moving on

green <- select(pomp, contains("green"))

green <- green %>% 
  mutate(sum = rowSums(green, na.rm = FALSE)) %>% 
  mutate(avg_green = sum/5) %>% 
  select(-sum)

comp <- select(pomp, contains("comp"))

comp <- comp %>% 
  mutate(sum = rowSums(comp, na.rm = FALSE)) %>% 
  mutate(avg_comp = sum/10) %>% 
  select(-sum)
  
intel <- select(pomp, contains("intel"))

intel <- intel %>% 
  mutate(sum = rowSums(intel, na.rm = FALSE)) %>% 
  mutate(avg_intel = sum/10) %>% 
  select(-sum)

open <- select(pomp, contains("open"))

open <- open %>% 
  mutate(sum = rowSums(open, na.rm = FALSE)) %>% 
  mutate(avg_open = sum/10) %>% 
  select(-sum)

avgs <- bind_cols(open, green, comp, intel) %>% 
  select(contains("avg"))

```
     
  4. Make plots that illustrate the distributions of the 4 POMP-scored variables.
  
```{r}

greenHistoPlot <- avgs %>% 
  ggplot() +
  aes(x = avgs$avg_green) +
  geom_histogram(fill = "orange")

greenHistoPlot 

compHistoPlot <- avgs %>% 
  ggplot() +
  aes(x = avgs$avg_comp) +
  geom_histogram(fill = "blue")

compHistoPlot

openHistoPlot <- avgs %>% 
  ggplot() +
  aes(x = avgs$avg_open) +
  geom_histogram(fill = "pink")

openHistoPlot

intelHistoPlot <- avgs %>% 
  ggplot() +
  aes(x = avgs$avg_intel) +
  geom_histogram(fill = "red")

intelHistoPlot

wrap_plots(intelHistoPlot, openHistoPlot, greenHistoPlot, compHistoPlot)
```
  5. Make scatterplots showing the relationships between **green reputation** and each personality trait.
    Include trend lines for **students** and **non-students**.
     What do these plots show?
```{r}

green_data$student <-
  recode_factor(green_data$student,
                "1" = "not a student",
                "2" = "student")

#didn't include the student column in my new data frame, so I guess I will merge the data frames to get the student column in there
avgs <- avgs %>% 
  mutate(id = c(1:263)) #creating an "id" column to merge them by

green_data <- green_data %>% 
  select(-id) #I want the id numbers to be 1:373 rather than the numbers the data came with

green_data <- green_data %>% 
  mutate(id = c(1:263))
           
avgs <- avgs %>% 
  full_join(green_data, by = "id")

avgs <- avgs %>% 
  select(id, student.x, contains("avg"))

avgs <- na.omit(avgs)
green_data <- na.omit(green_data)

#I'm sure there was a better way to get the "student" column into my "avgs" data frame, but at least what I did seemed to work

green_open <- avgs %>% 
  ggplot(aes(x = avg_green, y = avg_open, color = student.x)) + 
  geom_point(alpha = .5, color = "blue", fill = "blue") +
  geom_smooth(aes(group = student.x), se = FALSE) +
  theme(legend.position = "none") +
  scale_y_continuous("Openness to Experience", limits = c(30,100)) +
  scale_x_continuous("Green Reputation")
  

green_open

green_intel <- avgs %>% 
  ggplot(aes(x = avg_green, y = avg_intel, color = student.x)) +
  geom_point(alpha = .5,
             color = "blue",
             fill = "blue") +
  geom_smooth(aes(group = student.x), se = FALSE) +
  scale_y_continuous("Intellectual Curiosity", limits = c(30,100)) +
  scale_x_continuous("Green Reputation") +
  theme(legend.position = "top",
        legend.title = element_text(size = 5), 
        legend.text = element_text(size = 5))

green_intel

green_comp <- avgs %>% 
  ggplot(aes(x = avg_green,
             y = avg_comp,
             color = student.x)) +
  geom_point(alpha = .5,
             color = "blue",
             fill = "blue") +
  geom_smooth(aes(group = student.x), se = FALSE) +
  scale_y_continuous("Compassion", limits = c(30,100)) +
  scale_x_continuous("Green Reputation") +
  theme(legend.position = "none")

green_comp

plotplotplot <- wrap_plots(green_comp, green_open, green_intel)

plotplotplot

ggsave("plotplotplot.png", height = 6, width = 9) #save 

#getwd() to see file path (work directory)
#setwd() to give the path you want 
```
  
     Include trend lines for **students** and **non-students**.
     What do these plots show?
```{r} 
# There is a similar relationship between Green Reputation and Compassion for students and non-students.
# There is a positive correlation between Green Reputation and Openness to Experience for students, but there is no visible correlation between Green Reputation and Openness for non-students.
#Intellectual curiosity has a strong positive correlation with Green Reputation for students, but for non-students there is a weaker negative correlation. 
```
  6. Compare **green reputation** for students and non-students using a **rainfall plot** (bar + density + data points).
  
```{r}
ggplot(avgs) +
  aes(x = avg_green, 
      y = student.x, 
      color = student.x,
      fill = student.x) +
  geom_jitter(width = 0, height = .15, alpha = .3)+
  geom_boxplot(color = "black",
               alpha = .3,
               width = .2,
               size = .5) +
  ggdist::stat_slab(color = "black", 
    height = .3, 
    alpha = .5) + 
  xlab("Green Reputation") +
  theme(legend.position = "none") +
  theme_minimal()

```
  7. Compute a summary table of means, SDs, medians, minima, and maxima for the four total scores for students and non-students.
```{r}
summaryTable <- avgs %>% 
  group_by(student.x) %>% 
  summarize(across(avg_open:avg_intel,
      list(mean = ~mean(.x),
           sd = ~ sd(.x),
           min = ~min(.x),
           max = ~max(.x))))

summaryTable
```
  
In your assignment, prepare an RMarkdown file that includes both the coding and (for 4-7) narrative descriptions of your results.
