# Homework 03

You will analyze data looking at the relationship between **green reputation** and three personality traits--**compassion**, **intellectual curiosity**, and **openness to experiences**.
The dataset includes data from **students** and **non-students**.

```{r}
dictionary <- readr::read_csv("green_dictionary.csv")
green_data <- readr::read_csv("green_data.csv")

library(dplyr)
library(ggplot2)
library(tidyverse)
library(psychTools)

```

```{r}
view(dictionary)
view(green_data)
```
 
 ## 1. Inspect the item responses (e.g., with graphs or by summarizing distinct values). Is anything unusual?

```{r inspection}
green_data %>% 
  summary()
#What's unusual here is that minimum values are "-99" for several of these items. I'm going to change the "-99" values to "NA."
```
```{r}
 
my_green_data <- green_data %>% 
  mutate(across(everything(), 
                ~ na_if(.x, -99)))

view(my_green_data)
#Now there are no more "-99"s in my data.

```
## 2. Compute total scores for the four scales. Recode variables as needed.

```{r}
reversed <- dictionary %>% 
  filter(Keying < 0)
#making a new variable that filters those items that have a Keying less than 0 out of the dictionary. I should get 14 reverse coded variables.

view(reversed)
#looks good

#now I need to reverse code them
green_data_r <- my_green_data %>% 
  mutate(
    across(all_of(reversed),
           ~ recode(.x, "1" = 5, "2" = 4, "3" = 3, "4" = 4, "5" = 1),
           .names = "{.col}_r")
  )
#I keep getting error messages here!
#**just keep swimming**

#It looks like @samanthasheppard had a similar issue, so I'm going to borrow her solution so I can move forward until I figure out a better way. 
view(reversed)

green_data_r <- my_green_data %>%
  mutate(
    green2_r = recode(green2, "1" = "5", "2" = "4", "3" = "3", "4" = "2", "5" = "1"),
    comp6_r = recode(comp6, "1" = "5", "2" = "4", "3" = "3", "4" = "2", "5" = "1"),
    comp7_r = recode(comp7, "1" = "5", "2" = "4", "3" = "3", "4" = "2", "5" = "1"),
    comp8_r = recode(comp8, "1" = "5", "2" = "4", "3" = "3", "4" = "2", "5" = "1"),
    comp9_r = recode(comp9, "1" = "5", "2" = "4", "3" = "3", "4" = "2", "5" = "1"),
    comp10_r = recode(comp10, "1" = "5", "2" = "4", "3" = "3", "4" = "2", "5" = "1"),
    intel7_r = recode(intel7, "1" = "5", "2" = "4", "3" = "3", "4" = "2", "5" = "1"),
    intel8_r = recode(intel8, "1" = "5", "2" = "4", "3" = "3", "4" = "2", "5" = "1"),
    intel9_r = recode(intel9, "1" = "5", "2" = "4", "3" = "3", "4" = "2", "5" = "1"),
    intel10_r = recode(intel10, "1" = "5", "2" = "4", "3" = "3", "4" = "2", "5" = "1"),
    open7_r = recode(open7, "1" = "5", "2" = "4", "3" = "3", "4" = "2", "5" = "1"),
    open8_r = recode(open8, "1" = "5", "2" = "4", "3" = "3", "4" = "2", "5" = "1"),
    open9_r = recode(open9, "1" = "5", "2" = "4", "3" = "3", "4" = "2", "5" = "1"),     
    open10_r = recode(open10, "1" = "5", "2" = "4", "3" = "3", "4" = "2", "5" = "1")
  )
  
view(green_data_r)
#now I see the new variables coded properly but I can't find my new variable "open10_r" which is last on the list I just made. Why? 

#let's compute total scores for each person

new_green_data_r <- green_data_r %>% 
  rowwise() %>% 
  mutate(
    .id = id,
    green_total = sum(c_across(green1:green5), na.rm = TRUE),
    comp_total = sum(c_across(comp1:comp10), na.rm = TRUE), 
    intel_total = sum(c_across(intel1:intel10), na.rm = TRUE),
    open_total = sum(c_across(open1:open10), na.rm = TRUE)
  )

view(new_green_data_r)
#I'm not seeing my new totals variables. Don't know what went wrong there

```

## 3. Rescale the variables so that they go from 0-100 instead of the original range.
     Name the recaled variables `*_pomp`.
```{r} 
new_green_data_r <- new_green_data_r %>% 
  mutate(across(c(green1:open9_r &!student), 
                ~recode(.x, "1" = 0, "2" = 25, "3" = 50, "4" = 75, "5" = 100), 
                .names = "{.col}_pomp"))

view(new_green_data_r)
#why didn't that work? they look the same as before
```
     
  4. Make plots that illustrate the distributions of the 4 POMP-scored variables.
  5. Make scatterplots showing the relationships between **green reputation** and each personality trait.
     Include trend lines for **students** and **non-students**.
     What do these plots show?
  6. Compare **green reputation** for students and non-students using a **rainfall plot** (bar + density + data points).
  7. Compute a summary table of means, SDs, medians, minima, and maxima for the four total scores for students and non-students.
  
In your assignment, prepare an RMarkdown file that includes both the coding and (for 4-7) narrative descriptions of your results.
