---
title: "CLASS_PRACTICE"
author: "Antonietta"
date: "2/9/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(psych)
psych::bfi
```

1. Use the `psych::bfi` dataset.
2. Compute mean scores for each of the Big Five traits for each person.
3. Keep just Big Five scores, gender, age, and education.
4. Summarize the mean, standard deviation, minimum, and maximum values for the Big Five scores separately by gender groups.
5. Summarize the mean, standard deviation, minimum, and maximum values for the Big Five scores separately by educational level groups.

```{r, mutate and mean scores}
# Computing each mean score per person 

bfi_scored <- psych::bfi %>% 
  rowwise() %>%
  mutate(agree = mean(c_across(A1:A5), na.rm = TRUE), 
         con = mean(c_across(C1:C5), na.rm =  TRUE),
         extra = mean(c_across(E1:E5), na.rm = TRUE),
         neuro = mean(c_across(N1:N5), na.rm = TRUE), 
         open = mean(c_across(O1:O5), na.rm = TRUE), )%>%
    select(gender, education, age, agree, con, extra, neuro, open) 
  
# Sumarize by gender
  
bfi_scored %>% 
  group_by(gender) %>%
  summarize(across(agree:open, 
                   list(mean = mean , sd = sd , max = max , min = min)
                   )
            ) 


  # Sumarize by education 

bfi_scored %>%  
  group_by(education) %>%
  summarize(across(agree:open,
                   list(mean = mean , sd = sd , max = max , min = min)
                   )
            )
  

```

## rowwise tells r to do things differently by row, trating each row as its own group
## group_by separates data by group variable 
# head will return the table you made 


```{r}


```
 
 
 
 
 
 
 
 